Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} model system

### 2014-04-25 trunk/paper_enahpc14/Sec4:Experimental-results.tex
Manuel's changes

In  Figure~\ref{fig:3}, we  compare both  TTS (right  Y-axis)  and ETS
(left  Y-axis)  metrics  on  both  platforms.  As  expected,  Xeon  E5
outperforms Ivy Bridge EP, being  roughly 1.3x faster.  The reason for
that is  two-fold: \emph{i}) it has  higher clock frequency  than Ivy Bridge
(2.6  GHz  against  2.2 GHz),  and  \emph{ii})  it  aims at  computing  speed
regardless to  energy consumption.  In our experiments,  Ivy Bridge EP
showed the best energy-to-solution, reducing the energy consumption of
Xeon E5 by approximately 7\,\%.

@@

\subsection{Power-performance tracing on \tintorrum}
\label{subsec:4.3}

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution. First we run \cosmoart for a full-day simulation
and analyze the power consumption. Second we correlate performance with
power traces of two 1-hor time frames during the day: mid-day (from 11h-12h) and
midnight (23h-24h). All the experiments were performed using 192 \cosmoart processes
mapping 12 processes in each of node.

In order to obtain an initial overview of the power dissipation of the system model, we have 
obtained the power profiles using our \pmlib framework of a full-day simulation of \cosmoart 
leveraging the polling and blocking MPI policies, these results are depicted in the top/bottom 
plots of Figure~\ref{fig:9}, respectvely. o, captured with our PDUs. The first observation for 
the polling MPI policy is the quite plain power profile it generates. However, the first and 
last hours of the full-day simulation dissipate slightly less power than the mid-day hours. We 
relate these variations due to the chemical reactions and aerosols, computed by ART, that are 
taking place during the daylight hours but not during night. In this case peak power 
dissipation is about 3821\,W. Looking at the power profile using the blocking MPI policy, we 
can easily observe that the pattern is repeated favoring the reduction of power during 
nighttime hours in a more  noticeable way. The peak power dissipation in this mode is 3768\,W, 
which means a reduction of 50\,W in total or, in other words, 1,\% of reductuon in the peak 
power. Also, we observe power  drops in each hour simulated, basically related to periods in 
which the processes were performing blocking-waits instead of polling-waits, thus yielding the 
CPU to a low usage during this phases. Our final observation is the reduction in execution time 
when the blocking MPI policy is leveraged in contrast with the polling MPI. One can expect that 
the costs of the interruptions used in the blocking can lead to an increase of the total 
execution time, however, applications with extensive computing may show better perforamnce with 
the blocking mode. This plays in favour for our results since the energy consumed is also less 
by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

@@

\begin{figure*}[htbf]
  \centering
  \includegraphics[width=0.78\textwidth]{Figs/23_24_blq1_stat1.eps}
  \caption{blabla}
  \label{fig:8}
\end{figure*}

\begin{figure*}[htbf]
  \centering
  \scalebox{0.5}{\input{Figs/24hour.pstex_t}}
  \caption{24 hours simulation trace using the MPI blocking mode on
    \textsc{Tintorrum}.}
  \label{fig:9}
\end{figure*}

\begin{figure*}[htbf]
  \centering
  \scriptsize
  Simulation of 1 hour during midday (from 11h to 12h) leveraging the polling MPI policy\\
  \scalebox{0.5}{\input{Figs/11_12_blq0_figure.pstex_t}}\\
  Simulation of 1 hour during midday (from 11h to 12h) leveraging the blocking MPI policy\\
  \scalebox{0.5}{\input{Figs/11_12_blq1_figure.pstex_t}}\\
  Simulation of 1 hour during midnight (from 23h to 24h) leveraging the polling MPI policy\\
  \scalebox{0.5}{\input{Figs/23_24_blq0_figure.pstex_t}}\\
  Simulation of 1 hour during midnight (from 23h to 24h) leveraging the blocking MPI policy\\
  \scalebox{0.5}{\input{Figs/23_24_blq1_figure.pstex_t}}\\
  \begin{tabular}{rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rl}
& &  & &  & &  & &  \\[-0.15cm]
Computation     & \multicolumn{1}{>{\columncolor[RGB]{  0,  , 255}}p{0.4cm}}{}  & & 
Waiting a msg.  & \multicolumn{1}{>{\columncolor[RGB]{255,  0  ,0}}p{0.4cm}}{}  & & 
Block. send     & \multicolumn{1}{>{\columncolor[RGB]{255,  0,174}}p{0.4cm}}{}  & & 
Synchronization & \multicolumn{1}{>{\columncolor[RGB]{179,  0,  0}}p{0.4cm}}{}  & & 
Wait/Wait all   & \multicolumn{1}{>{\columncolor[RGB]{235,  0,  0}}p{0.4cm}}{}  \\
& &  & &  & &  & &  \\[-0.15cm]
Immediate recv. & \multicolumn{1}{>{\columncolor[RGB]{100,100,177}}p{0.4cm}}{}  & &
I/O             & \multicolumn{1}{>{\columncolor[RGB]{172,174, 41}}p{0.4cm}}{}  & &
Group comm.     & \multicolumn{1}{>{\columncolor[RGB]{255,144, 26}}p{0.4cm}}{}  & &
Others          & \multicolumn{1}{>{\columncolor[RGB]{192,224,  0}}p{0.4cm}}{}  & &
Idle            & \multicolumn{1}{>{\columncolor[RGB]{117,195,255}}p{0.4cm}}{}  \\
\end{tabular}
  \caption{Power-perfomance traces during midday and midnight leveraging the polling and blocking MPI policies on \tinto.}
  \label{fig:11}
\end{figure*}

>>>>

In  Figure~\ref{fig:3}, we  compare both  TTS (right  Y-axis)  and ETS
(left  Y-axis)  metrics  on  both  platforms.  As  expected,  Xeon  E5
outperforms Ivy Bridge EP, being  roughly 1.3x faster.  The reason for
that is  two-fold: \emph{i}) it has  higher clock frequency  than Ivy Bridge
(2.6\,GHz  against  2.2\,GHz),  and  \emph{ii})  it  aims at  computing  speed
regardless to  energy consumption.  In our experiments,  Ivy Bridge EP
showed the best energy-to-solution, reducing the energy consumption of
Xeon E5 by approximately 7\,\%.

@@

\subsection{Power-performance tracing on \tinto}
\label{subsec:4.3}

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution using the attached PDUs in each node. First we run \cosmoart 
for a full-day simulation and analyze the power consumption. Second we correlate performance 
with power traces of two 1-hor time frames during the day: mid-day (from 11h-12h) and
midnight (23h-24h). All the experiments were performed using 192 \cosmoart processes
mapping 12 processes in each of node.

% For all the experiments we also analyze the contribution of the MPI library to the energy 
% consumption and how the use of the blocking versus polling message-passing policies can 
% potentially render energy savings. Specifically, the OpenMPI 1.6.5 library installed in \tinto, 
% features two operation modes, blocking and modes, which can be selected before \cosmoart is 
% launched. In the polling configuration (the detault mode) ``MPI'' continually polls
% the netwrok interface to check for the completion of an event (e.g. send or receive).
% This usually yields low latency but high CPU utilization. Thus, can one expcet that this mode attanins the best performance, possibley at the cost a higher energy usage. In the blocking
% mode, the CPU is yield to other processes/threads if there are no incoming messages.

In order to obtain an initial overview of the power dissipation of the system model, we have 
obtained the power profiles using our \pmlib framework of a full-day simulation of \cosmoart 
leveraging the polling and blocking MPI policies, these results are depicted in the top/bottom 
plots of Figure~\ref{fig:9}, respectvely. The first observation for 
the polling MPI policy is the quite plain power profile it generates. However, the first and 
last hours of the full-day simulation dissipate slightly less power than the mid-day hours. We 
relate these variations due to the chemical reactions and aerosols, computed by ART, that are 
taking place during the daylight hours but not during night. In this case peak power 
dissipation is about 3821\,W. Looking at the power profile using the blocking MPI policy, we 
can easily observe that the pattern is repeated favoring the reduction of power during 
nighttime hours in a more  noticeable way. The peak power dissipation in this mode is 3768\,W, 
which means a reduction of 50\,W in total or, in other words, 1,\% of reductuon in the peak 
power. Also, we observe power  drops in each hour simulated, basically related to periods in 
which the processes were performing blocking-waits instead of polling-waits, thus yielding the 
CPU to a low usage during this phases. Our final observation is the reduction in execution time 
when the blocking MPI policy is leveraged in contrast with the polling MPI. One can expect that 
the costs of the interruptions used in the blocking can lead to an increase of the total 
execution time, however, applications with extensive computing may show better perforamnce with 
the blocking mode. This plays in favour for our results since the energy consumed is also less 
by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

@@

% \begin{figure*}[htbf]
%   \centering
%   Simulation of 1 hour during midnight (from 23h to 24h) leveraging the polling MPI policy\\
%   \includegraphics[width=0.78\textwidth]{Figs/23_24_blq0_stat.eps}\\
%   Percentage of time for the 1 hour during midnight (from 23h to 24h) leveraging the polling MPI policy\\
%   \includegraphics[width=0.78\textwidth]{Figs/23_24_blq1_stat.eps}\\
%   \caption{blabla}
%   \label{fig:8}
% \end{figure*}

\begin{figure*}[htbf]
  \centering
  \hspace{0.8cm}
  \scalebox{0.5}{\input{Figs/24hour.pstex_t}}
  \caption{24 hours simulation trace using the MPI blocking mode on \tinto.}
  \label{fig:9}
\end{figure*}

\begin{figure*}[htbf]
  \centering
  \scriptsize
  Simulation of 1 hour during midday (from 11h to 12h) leveraging the polling MPI policy\\
  \scalebox{0.5}{\input{Figs/11_12_blq0_figure.pstex_t}}\\
  Simulation of 1 hour during midday (from 11h to 12h) leveraging the blocking MPI policy\\
  \scalebox{0.5}{\input{Figs/11_12_blq1_figure.pstex_t}}\\
  1 hour simulation during midnight (from 23h to 24h) leveraging the polling MPI policy\\
  \scalebox{0.5}{\input{Figs/23_24_blq0_figure.pstex_t}}\\
  1 hour simulation during midnight (from 23h to 24h) leveraging the blocking MPI policy\\
  \scalebox{0.5}{\input{Figs/23_24_blq1_figure.pstex_t}}\\
  \begin{tabular}{rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rl}
& &  & &  & &  & &  \\[-0.15cm]
Computation     & \multicolumn{1}{>{\columncolor[RGB]{  0,  , 255}}p{0.4cm}}{}  & & 
Waiting a msg.  & \multicolumn{1}{>{\columncolor[RGB]{255,  0  ,0}}p{0.4cm}}{}  & & 
Block. send     & \multicolumn{1}{>{\columncolor[RGB]{255,  0,174}}p{0.4cm}}{}  & & 
Synchronization & \multicolumn{1}{>{\columncolor[RGB]{179,  0,  0}}p{0.4cm}}{}  & & 
Wait/Wait all   & \multicolumn{1}{>{\columncolor[RGB]{235,  0,  0}}p{0.4cm}}{}  \\
& &  & &  & &  & &  \\[-0.15cm]
Immediate recv. & \multicolumn{1}{>{\columncolor[RGB]{100,100,177}}p{0.4cm}}{}  & &
I/O             & \multicolumn{1}{>{\columncolor[RGB]{172,174, 41}}p{0.4cm}}{}  & &
Group comm.     & \multicolumn{1}{>{\columncolor[RGB]{255,144, 26}}p{0.4cm}}{}  & &
Others          & \multicolumn{1}{>{\columncolor[RGB]{192,224,  0}}p{0.4cm}}{}  & &
Idle            & \multicolumn{1}{>{\columncolor[RGB]{117,195,255}}p{0.4cm}}{}  \\
\end{tabular}
  \caption{Power-perfomance traces during midday and midnight leveraging the polling and blocking MPI policies on \tinto.}
  \label{fig:11}
\end{figure*}
